---
title: "Exercise 1 - Linear models for Gaussian data"
author: Anders Christiansen Sørby, Edvard Hove
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("car")
install.packages("GGally")
library(car)
library(GGally)
library(mylm)
```

We will work on the dataset `carData`, which consists of 3987 observations on the following 5 variables:

* `wages`, composite hourly wage rate from all jobs
* `education`, number of years in schooling
* `age`, in years
* `sex`, Male or Female
* `language`, English, French or Other

```{r cars, echo=FALSE, message=FALSE}
data(SLID, package = "carData")
SLID <- SLID[complete.cases(SLID), ]
```

Using the function `ggpairs` we get the following diagnostic plot matrix:

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE}
ggpSLID = ggpairs(SLID)
print(ggpSLID,progress=F)
```
The plot suggests that `wages` increase with both `education` and `age`, despite the apparent slight decrease of `education` with `age`. It also seems like males recieve higher wages than females, despite similar levels of age and education. The `language` variable seems to have some relationship with both `language` and `education` 
