---
title: "Linear models for Gaussian data"
subtitle: "TMA4315 - Exercise 1"
author: Anders Christiansen Sørby, Edvard Hove
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("car")
#install.packages("GGally")
library(car)
library(GGally)
library(mylm)
```

We will work on the dataset `carData`, which consists of 3987 observations on the following 5 variables:

* `wages`, composite hourly wage rate from all jobs
* `education`, number of years in schooling
* `age`, in years
* `sex`, Male or Female
* `language`, English, French or Other

```{r cars, echo=FALSE, message=FALSE}
data(SLID, package = "carData")
SLID <- SLID[complete.cases(SLID), ]
```

Using the function `ggpairs` we get the following diagnostic plot matrix:

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE}
ggpSLID = ggpairs(SLID)
print(ggpSLID,progress=F)
```
The plot suggests that `wages` increase with both `education` and `age`, despite the apparent slight decrease of `education` with `age`. It also seems like males recieve higher wages than females, despite similar levels of age and education. There seems to be some relationship between `language` and `age`, but it is not clear if there is a corresponding relationship between `language` and `wages`. 

When performing a multiple linear regression analysis to study how `wages` depends on the explanatory variables we need to assume that: ELABORATE

- each data point \(Y_i\) is independent
- the residuals are normally distributed with a homogenous variance, such that \(Y_i \sim \mathcal{N}(\mu_i,\sigma^2)\)
- there is a linear relationship between `wages` and the explanatory variables, such that \(\mu_i = \eta_i = \mathbf{x_i}^T\beta\)
```{r eval=FALSE FILE="R/mylm.R"}
```
