---
title: "Linear models for Gaussian data"
author: "Anders Christiansen Sørby, Edvard Hove, Angela Maiken Johnsen"
output:
  pdf_document: default
#  html_document: default
subtitle: TMA4315 - Exercise 1
header-includes: \usepackage{nicefrac}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("car")
#install.packages("GGally")
library(car)
library(GGally)
library(mylm)
```

# Part 1: Explanatory analysis

We will work on the dataset `carData`, which consists of 3987 observations on the following 5 variables:

* `wages`, composite hourly wage rate from all jobs
* `education`, number of years in schooling
* `age`, in years
* `sex`, Male or Female
* `language`, English, French or Other

```{r cars, echo=FALSE, message=FALSE}
data(SLID, package = "carData")
SLID <- SLID[complete.cases(SLID), ]
n <- length(SLID[,1])
```

Using the function `ggpairs` we get the following diagnostic plot matrix:

```{r pressure, echo=FALSE, message=FALSE, warning=FALSE}
ggpSLID = ggpairs(SLID)
print(ggpSLID,progress=F)
```
The plot suggests that `wages` increase with both `education` and `age`, despite the apparent slight decrease of `education` with `age`. It also seems like males recieve higher wages than females, despite similar levels of age and education. There seems to be some relationship between `language` and `age`, but it is not clear if there is a corresponding relationship between `language` and `wages`. 

When performing a multiple linear regression analysis to study how `wages` depends on the explanatory variables we need to assume that: ELABORATE

- each data point \(Y_i\) is independent
- the residuals are normally distributed with a homogenous variance, such that \(Y_i \sim \mathcal{N}(\mu_i,\sigma^2)\)
- there is a linear relationship between `wages` and the explanatory variables, such that \(\mu_i = \eta_i = \mathbf{x_i}^T\beta\)


# Part 2: Simple linear regression with the mylm package

**a)**

```{r model1}
model1 <- mylm(wages ~education, data = SLID)
print(model1)

model1b <- lm(wages ~ education, data = SLID)
print(model1b)
```

**b)**

```{r summary}
summary(model1)
#summary(model1b)
```


**c)**
```{r plot}
plot(model1)
```

**d)**

The error of the regression can be represented using the following quantities. This holds for linear regression. We are using the following naming convention, which may differ from other texts.

\begin{align}
SSE = \sum^{n}_{i=1}(\hat{y}_i - \bar{y})^2 \\
SSR = \sum^{n}_{i=1}(y_i - \hat{y}_i)^2 \\
SST = \sum^{n}_{i=1}(y_i - \bar{y}_i)^2 = SSE + SSR \\
R^2 = 1 - \frac{SSR}{SST}
\end{align}
where $\bar{y}$ is the mean of responses and $\hat{y}_i$ is the fitted value. 

The residual sum of squares (SSR) for this model is \(`r sprintf('%.0f', model1$ssr)`\). 
The model has \(`r n-model1$p`\) degrees of freedom, which is obtained by subtracting the number of estimated parameters from the number of observations.
The total sum of squares (SST) is \(`r sprintf('%.0f', model1$sst)`\). 
For a model with so many degrees of freedom we can use the limiting distribution

\begin{equation}
(p-1)F = (p-1)\frac{\nicefrac{SSE}{p-1}}{\nicefrac{SSR}{n-p}} \sim \chi_{p-1}^2,
\end{equation}

which gives gives a value of \(`r sprintf('%.2f', (n-model1$p)*model1$sse/model1$ssr)`\).
For a $\chi_{1}^2$ distribution, where the mean is 1, this is a very extreme value.
Hence the regression is clearly significant.
A $\chi^2_k$ distribution is the sum of the squares of $k$ independent standard normal random variables.
In our case we have $k=1$, which means that the $\chi^2$-statistic is simply the square of the $z$-statistic. 
The critical value for a $\chi_{1}^2$ test with confidence of 95% is \(`r sprintf('%.2f',qchisq(0.95,1))`\), while the two-sided $z$-test has critical values $\pm 1.96$.


**e)**
The coefficient of determination, \(R^2\), shows how much of the variance of the prediction can be explained by the variance of the response.
This gives some information about models goodness of fit.
In our case \(R^2\approx0.09\), which indicates that considering a model with additional covariates is a good idea.
A high proportion of the variance in the data is not accounted for by `model1`.

# Part 3: Multiple linear regression

**a)**

```{r}
model2 <- mylm(wages ~ education + age, data = SLID)
summary(model2)
```

**b)**

From the summary above we see that the intercept and both the regression coefficients are significantly different from zero according to the $z$-test. 

**c)**
```{r}
model2c1 <- mylm(wages ~ education, data = SLID)
model2c2 <- mylm(wages ~ age, data = SLID)
summary(model2c1)
summary(model2c2)
```
The values differ from the bigger model because each covariate is used to explain more of the resonse than in the bigger model. 
We know from the plot in part 1 that `age` and `education` are correlated.
This means that changing the covariate in one of the small models is similar to changing both covariates in the bigger model.
For uncorrelated covariates we would expect the same parameter estimates from the simple and multiple linear regression.

# Part 4: Testing mylm

Now we can test our function mylm on any linear model.

```{r}
model4a <- mylm(wages ~ sex + age + language + education^2, data = SLID)
model4b <- mylm(wages ~ age + language + language * age, data = SLID)
model4c <- mylm(wages ~ education -1, data = SLID)
summary(model4a)
summary(model4b)
summary(model4c)
```
